Taper Models

### 2021-09-29 analysis/get_util_piece_size.Rmd
old FMUs in Alpac and Millar FMAs merged.  some piece size work done

```{r libraries, echo=FALSE, message=FALSE}
#setwd("/home/gwa/Work/historical-residue/Analysis")
setwd("g:/Shared drives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/Analysis")

library(tidyverse)
# library(ggthemes)
# library(gridExtra)
# library(viridis)
# library(skimr)
# library(scales)
# library(DescTools)
# library(zoo)
# library(ezknitr)
#ezknit(file="analysis/manuscript.Rmd", out_dir="output", fig_dir="myfigs")
```

@@

```{r read data files, echo=FALSE}
util_std <- read.csv(file = '../data/util_std.csv')
piece_size <- read.csv(file = '../data/piece_size.csv')
fmu_twp <- read.csv(file ='../data/FMU_TWP.csv')
head(util_std)
head(piece_size)
head(fmu_twp)
util_piece <-full_join(util_std,piece_size,by=c("FMU","CD"))
util_piece_fmu_twp <- full_join(util_piece,fmu_twp,by="FMU")
util_piece_fmu_twp$piece_size <- 1/util_piece_fmu_twp$logspm3
util_piece_fmu_twp

```

>>>>

```{r libraries, echo=FALSE, message=FALSE}
#setwd("/home/gwa/Work/historical-residue/Analysis")
setwd("/mnt/chromeos/GoogleDrive/SharedDrives/FES_Data_Sharing_Forestry/Manuscripts/histressupp/analysis")

library(tidyverse)
library(janitor)
# library(ggthemes)
# library(gridExtra)
# library(viridis)
# library(skimr)
# library(scales)
# library(DescTools)
# library(zoo)
# library(ezknitr)
#ezknit(file="analysis/manuscript.Rmd", out_dir="output", fig_dir="myfigs")
```

@@

```{r read data files, echo=FALSE}
util_std <- read.csv(file = '../data/util_std.csv')
piece_size <- read.csv(file = '../data/piece_size.csv')
fmu_twp <- read.csv(file ='../data/FMU_TWP.csv')
species_prop <- read.csv(file = '../data/species_proportions_full.csv') %>%
  select(TRM,max_species)

fmu_twp_c <- fmu_twp
fmu_twp_d <- fmu_twp

fmu_twp_c$CD <- "Conif"
fmu_twp_d$CD <- "Decid"
fmu_twp <- rbind(fmu_twp_c,fmu_twp_d)

#head(util_std)
#head(piece_size)
#head(fmu_twp)

util_piece <-full_join(util_std,piece_size,by=c("FMU","CD"))
util_piece_fmu_twp <- full_join(util_piece,fmu_twp,by=c("FMU","CD"))
util_piece_fmu_twp$piece_size <- 1/util_piece_fmu_twp$logspm3
util_piece_fmu_twp_species <- full_join(util_piece_fmu_twp,species_prop,by="TRM")
util_piece_fmu_twp_species <- subset(util_piece_fmu_twp_species, (FMU != "A14") &(FMU != "A15") &(FMU != "W11")& (FMU != "W13"))

subs <- subset(util_piece_fmu_twp, (!is.na(piece_size)) & (!is.na(areaha)) & (CD == "Decid"))
avepieceD <- weighted.mean(subs$piece_size,subs$areaha,na.rm = TRUE)

subs <- subset(util_piece_fmu_twp, (!is.na(piece_size)) & (!is.na(areaha)) & (CD == "Conif"))
avepieceC <- weighted.mean(subs$piece_size,subs$areaha,na.rm = TRUE)

avepieceC
avepieceD

head(util_piece_fmu_twp_species)
#if (util_piece_fmu_twp_species$CD == "Decid") util_piece_fmu_twp_species$max_species <- "popu_tre"
tabyl(util_piece_fmu_twp_species,FMU,CD,show_na = TRUE)
#if (util_piece_fmu_twp_species$CD == "Decid") util_piece_fmu_twp_species$max_species <- "popu_tre"
#if (util_piece_fmu_twp_species$max_species == "pseu_men") util_piece_fmu_twp_species$max_species <- "pice_gla"

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(max_species = case_when(
    CD == "Decid" ~ "popu_tre",
    (CD == "Conif") & (max_species == "popu_tre") ~ "pice_gla",
    (CD == "Conif") & (max_species == "popu_bal") ~ "pice_gla",
    (CD == "Conif") & (max_species == "pseu_men") ~ "pice_gla",
    (CD == "Conif") & (max_species == "") ~ "pice_gla",
    CD == "Conif" ~ max_species
    )
  )

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(piece_size = case_when(
    (CD == "Conif") & (is.na(piece_size)) ~ avepieceC,
    (CD == "Decid") & (is.na(piece_size)) ~ avepieceD,
    TRUE ~ piece_size
    )
  )

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(stumpht = case_when(
    (CD == "Conif") & (is.na(stumpht)) ~ 30.0,
    (CD == "Decid") & (is.na(stumpht)) ~ 30.0,
    TRUE ~ stumpht
    )
  )

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(merch_len = case_when(
    (CD == "Conif") & (is.na(merch_len)) ~ 4.88,
    (CD == "Decid") & (is.na(merch_len)) ~ 4.88,
    TRUE ~ merch_len
    )
  )

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(stumpd = case_when(
    (CD == "Conif") & (is.na(stumpd)) ~ 15.0,
    (CD == "Decid") & (is.na(stumpd)) ~ 15.0,
    TRUE ~ stumpd
    )
  )

util_piece_fmu_twp_species <- util_piece_fmu_twp_species %>%
  mutate(topd = case_when(
    (CD == "Conif") & (is.na(topd)) ~ 11.0,
    (CD == "Decid") & (is.na(topd)) ~ 10.0,
    TRUE ~ topd
    )
  )

tabyl(util_piece_fmu_twp_species,max_species,CD,show_na = TRUE)
head(util_piece_fmu_twp_species)
util_piece_fmu_twp_species
```

### 2021-09-29 analysis/taper.Rmd
old FMUs in Alpac and Millar FMAs merged.  some piece size work done

diameters <- c(15,15.5,16,16.5,17,17.5,18,18.5,19,19.5,20,20.5,21,21.5,22,22.5,23,23.5,24,24.5,25,25.5,26,26.5,27,27.5,28,28.5,29,29.5,30
)
ecozones <- c(4,5,6,9,10,14)
speciesList <- c(",ABIE,LAS,,",",ABIE,BAL,,",",BETU,PAP,PAP,",",LARI,LAR,,",",PICE,ENG,,",",PICE,GLA,,",",PICE,MAR,,",
                 ",PINU,BAN,,",",PINU,CON,LAT,",",POPU,TRE,,",",POPU,BAL,,",",PSEU,MEN,GLA,")
species2 <- c("Fa","Fb","Bw","Lt","Se","Sw","Sb",
              "Pj","Pl","Aw","Pb","Fd")
header <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height"
header2 <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height,merchVol,topVolRatio,topBarkRatio"

cat(sprintf("%s\n",header),file = "../data/bioCalcInput.csv", append = FALSE)
cat(sprintf("%s\n",header2),file = "../data/ratios.csv", append = FALSE)

@@

treenum <- 0
spcnt <- 0
for (sp in species2) {
  spcnt <- spcnt + 1
  t <- taperCoefficients %>% filter(species == species2[[spcnt]])
  b0 <- t$b0
  b1 <- t$b1
  b2 <- t$b2
  topD <- t$uTopD
  stumpht <- t$uStumpHt
  minloglength <- t$uMinlen

@@

      cat(sprintf("%i%s%6.2f%s%6.2f\n",treenum,stuff,dbh,",",ht),
          file = "../data/bioCalcInput.csv",append = TRUE)
      cat(sprintf("%i%s%6.2f%s%6.2f%s%f%s%f%s%f\n",treenum,stuff,dbh,",",ht,",",
                  merchVol,",",topVolRatio,",",topBarkRatio),
          file = "../data/ratios.csv",append = TRUE)
    }
  } 
}
#closeAllConnections()
```

>>>>

diameters <- c(
13	,
14	,
15	,
16	,
17	,
18	,
19	,
20	,
21	,
22	,
23	,
24	,
25	,
26	,
27	,
28	,
29	,
30	,
31	,
32	,
33	,
34	,
35	,
36	,
37	,
38	,
39	,
40	,
41	,
42	,
43	,
44	,
45	,
46	,
47	,
48	,
49	,
50	
)

ecozones <- c(4,5,6,9,10,14)
speciesList <- c(",ABIE,LAS,,",",ABIE,BAL,,",",BETU,PAP,PAP,",",LARI,LAR,,",",PICE,ENG,,",",PICE,GLA,,",",PICE,MAR,,",
                 ",PINU,BAN,,",",PINU,CON,LAT,",",POPU,TRE,,",",POPU,BAL,,",",PSEU,MEN,GLA,")
species2 <- c("Fa","Fb","Bw","Lt","Se","Sw","Sb",
              "Pj","Pl","Aw","Pb","Fd")
header <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height"
header2 <- "Tree_num,Province,Ecozone,Genus,Species,Variety,DBH,Height,merchVol,topVolRatio,topBarkRatio"

cat(sprintf("%s\n",header),file = "e:/ualberta/bioCalcInput.csv", append = FALSE)
cat(sprintf("%s\n",header2),file = "../data/ratios.csv", append = FALSE)

@@

treenum <- 0
spcnt <- 0
for (sp in species2) {
  spcnt <- spcnt + 1
  t <- taperCoefficients %>% filter(species == species2[[spcnt]])
  b0 <- t$b0
  b1 <- t$b1
  b2 <- t$b2
  topD <- t$uTopD
#  stumpht <- t$uStumpHt
  stumpht <- 0
  minloglength <- t$uMinlen

@@

      cat(sprintf("%i%s%6.2f%s%6.2f\n",treenum,stuff,dbh,",",ht),
          file = "e:/ualberta/bioCalcInput.csv",append = TRUE)
      cat(sprintf("%i%s%6.2f%s%6.2f%s%f%s%f%s%f\n",treenum,stuff,dbh,",",ht,",",
                  merchVol,",",topVolRatio,",",topBarkRatio),
          file = "../data/ratios.csv",append = TRUE)
    }
  } 
}
#closeAllConnections()
```
